---
title: "faers_data_capstone"
author: "MK NYARKO"
date: "2025-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(dplyr)
library(stringr)
library(tidyr)
```

# Q1 2024 DATA MANAGEMENT ONLY

```{r}
# Paths
faers_path <- "/Users/mknyarko/Documents/COURSES BROWN SCHOOL/CAPSTONE/CAPSTONE PROJECT/ANALYSIS FILES/FAERS DATA 2024/faers_ascii_2024q1/ASCII"
```

```{r}
# Load DRUG table
drug_q1 <- fread(file.path(faers_path, "DRUG24Q1.txt"), sep = "$")
```

```{r}
# Extract semaglutide-related reports
semaglutide_q1 <- drug_q1[
  grepl("SEMAGLUTIDE|OZEMPIC|WEGOVY|RYBELSUS", drugname, ignore.case = TRUE) |
  grepl("SEMAGLUTIDE", prod_ai, ignore.case = TRUE)
]
```

```{r}
# Vector to keep only Semaglutide CASEIDs
semaglutide_ids_q1 <- unique(semaglutide_q1$caseid)
```

```{r}
# Load other Q1 tables
demo_q1 <- fread(file.path(faers_path, "DEMO24Q1.txt"), sep = "$")
reac_q1 <- fread(file.path(faers_path, "REAC24Q1.txt"), sep = "$")
outc_q1 <- fread(file.path(faers_path, "OUTC24Q1.txt"), sep = "$")
```

```{r}
# Restrict to semaglutide CASEIDs
  demo_q1_ids <- demo_q1[caseid %in% semaglutide_ids_q1]
  reac_q1_ids <- reac_q1[caseid %in% semaglutide_ids_q1]
  outc_q1_ids <- outc_q1[caseid %in% semaglutide_ids_q1]
```

```{r}
# Limit to U.S. cases only
  demo_q1_ids_US <- demo_q1_ids[reporter_country == "US"]
```

```{r}
# Merge: demographics + outcomes
  merged_demo_outc_q1 <- merge(demo_q1_ids_US, outc_q1_ids, by = "caseid", all.x = TRUE)
```

```{r}
# Merge in reactions
  merged_d_o_reac_q1 <- merge(merged_demo_outc_q1, reac_q1_ids, by = "caseid", all.x = TRUE)
```

```{r}
 # Merge in semaglutide drug info
  merged_all_q1 <- merge(merged_d_o_reac_q1, semaglutide_q1, by = "caseid", all.x = TRUE)
```

```{r}
# Make duplicate names unique
names(merged_all_q1) <- make.unique(names(merged_all_q1))

```


```{r}
# Remove duplicate columns
merged_all_q1_clean <- merged_all_q1 %>%
  select(-primaryid.x.1, -primaryid.y.1)
```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q1)

```


```{r}
# Vector to keep only Semaglutide CASEIDs
merged_all_q1_ids <- unique(merged_all_q1_clean$caseid)
```

```{r}
# Define severity ranking (custom order)
severity_order <- c("DE", "LT", "HO", "DS", "CA", "RI", "OT")

```

```{r}
# Clean dataset so that each CASEID has only the most serious outcome
merged_all_q1_serious <- merged_all_q1_clean %>%
  mutate(
    severity_rank = match(outc_cod, severity_order)  # assign rank
  ) %>%
  group_by(caseid) %>%
  slice_min(order_by = severity_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-severity_rank)  # drop helper column if not needed
```

```{r}
# Check results, compared raw merged to serious merged, then serious merged to reference unique case id
nrow(merged_all_q1)                     # original rows (with duplicates)
length(unique(merged_all_q1$caseid))    # number of unique CASEIDs
nrow(merged_all_q1_serious)             # should equal number of unique CASEIDs
```

```{r}
# Write the 2024 Q1 Full csv file
fwrite(merged_all_q1_serious, "semaglutide_2024Q1_full.csv")
```

# Q2 2024 DATA MANAGEMENT ONLY

```{r}
# Paths
faers_path2 <- "/Users/mknyarko/Documents/COURSES BROWN SCHOOL/CAPSTONE/CAPSTONE PROJECT/ANALYSIS FILES/FAERS DATA 2024/faers_ascii_2024q2/ASCII"
```

```{r}
# Load DRUG table
drug_q2 <- fread(file.path(faers_path2, "DRUG24Q2.txt"), sep = "$")
```

```{r}
# Extract semaglutide-related reports
semaglutide_q2 <- drug_q2[
  grepl("SEMAGLUTIDE|OZEMPIC|WEGOVY|RYBELSUS", drugname, ignore.case = TRUE) |
  grepl("SEMAGLUTIDE", prod_ai, ignore.case = TRUE)
]
```

```{r}
# Vector to keep only Semaglutide CASEIDs
semaglutide_ids_q2 <- unique(semaglutide_q2$caseid)
```

```{r}
# Load other Q2 tables
demo_q2 <- fread(file.path(faers_path2, "DEMO24Q2.txt"), sep = "$")
reac_q2 <- fread(file.path(faers_path2, "REAC24Q2.txt"), sep = "$")
outc_q2 <- fread(file.path(faers_path2, "OUTC24Q2.txt"), sep = "$")
```

```{r}
# Restrict to semaglutide CASEIDs
  demo_q2_ids <- demo_q2[caseid %in% semaglutide_ids_q2]
  reac_q2_ids <- reac_q2[caseid %in% semaglutide_ids_q2]
  outc_q2_ids <- outc_q2[caseid %in% semaglutide_ids_q2]
```

```{r}
# Limit to U.S. cases only
  demo_q2_ids_US <- demo_q2_ids[reporter_country == "US"]
```

```{r}
# Merge: demographics + outcomes
  merged_demo_outc_q2 <- merge(demo_q2_ids_US, outc_q2_ids, by = "caseid", all.x = TRUE)
```

```{r}
# Merge in reactions
  merged_d_o_reac_q2 <- merge(merged_demo_outc_q2, reac_q2_ids, by = "caseid", all.x = TRUE)
```

```{r}
 # Merge in semaglutide drug info
  merged_all_q2 <- merge(merged_d_o_reac_q2, semaglutide_q2, by = "caseid", all.x = TRUE)
```

```{r}
# Make duplicate names unique
names(merged_all_q2) <- make.unique(names(merged_all_q2))

```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q2)
```


```{r}
# Remove duplicate columns
merged_all_q2_clean <- merged_all_q2 %>%
  select(-primaryid.x.1, -primaryid.y.1)
```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q2_clean)
```


```{r}
# Vector to keep only Semaglutide CASEIDs
merged_all_q2_ids <- unique(merged_all_q2_clean$caseid)
```

```{r}
# Define severity ranking (custom order)
severity_order <- c("DE", "LT", "HO", "DS", "CA", "RI", "OT")

```

```{r}
# Clean dataset so that each CASEID has only the most serious outcome
merged_all_q2_serious <- merged_all_q2_clean %>%
  mutate(
    severity_rank = match(outc_cod, severity_order)  # assign rank
  ) %>%
  group_by(caseid) %>%
  slice_min(order_by = severity_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-severity_rank)  # drop helper column if not needed
```

```{r}
# Check results, compared raw merged to serious merged, then serious merged to reference unique case id
nrow(merged_all_q2)                     # original rows (with duplicates)
length(unique(merged_all_q2$caseid))    # number of unique CASEIDs
nrow(merged_all_q2_serious)             # should equal number of unique CASEIDs
```

```{r}
# Write the 2024 Q2 Full csv file
fwrite(merged_all_q2_serious, "semaglutide_2024Q2_full.csv")
```

# Q3 2024 DATA MANAGEMENT ONLY

```{r}
# Paths
faers_path3 <- "/Users/mknyarko/Documents/COURSES BROWN SCHOOL/CAPSTONE/CAPSTONE PROJECT/ANALYSIS FILES/FAERS DATA 2024/faers_ascii_2024q3/ASCII"
```

```{r}
# Load DRUG table
drug_q3 <- fread(file.path(faers_path3, "DRUG24Q3.txt"), sep = "$")
```

```{r}
# Extract semaglutide-related reports
semaglutide_q3 <- drug_q3[
  grepl("SEMAGLUTIDE|OZEMPIC|WEGOVY|RYBELSUS", drugname, ignore.case = TRUE) |
  grepl("SEMAGLUTIDE", prod_ai, ignore.case = TRUE)
]
```

```{r}
# Vector to keep only Semaglutide CASEIDs
semaglutide_ids_q3 <- unique(semaglutide_q3$caseid)
```

```{r}
# Load other Q3 tables
demo_q3 <- fread(file.path(faers_path3, "DEMO24Q3.txt"), sep = "$")
reac_q3 <- fread(file.path(faers_path3, "REAC24Q3.txt"), sep = "$")
outc_q3 <- fread(file.path(faers_path3, "OUTC24Q3.txt"), sep = "$")
```

```{r}
# Restrict to semaglutide CASEIDs
  demo_q3_ids <- demo_q3[caseid %in% semaglutide_ids_q3]
  reac_q3_ids <- reac_q3[caseid %in% semaglutide_ids_q3]
  outc_q3_ids <- outc_q3[caseid %in% semaglutide_ids_q3]
```

```{r}
# Limit to U.S. cases only
  demo_q3_ids_US <- demo_q3_ids[reporter_country == "US"]
```

```{r}
# Merge: demographics + outcomes
  merged_demo_outc_q3 <- merge(demo_q3_ids_US, outc_q3_ids, by = "caseid", all.x = TRUE)
```

```{r}
# Merge in reactions
  merged_d_o_reac_q3 <- merge(merged_demo_outc_q3, reac_q3_ids, by = "caseid", all.x = TRUE)
```

```{r}
 # Merge in semaglutide drug info
  merged_all_q3 <- merge(merged_d_o_reac_q3, semaglutide_q3, by = "caseid", all.x = TRUE)
```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q3)
```


```{r}
# Make duplicate names unique
names(merged_all_q3) <- make.unique(names(merged_all_q3))

```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q3)
```


```{r}
# Remove duplicate columns
merged_all_q3_clean <- merged_all_q3 %>%
  select(-primaryid.x.1, -primaryid.y.1)
```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q3_clean)
```


```{r}
# Vector to keep only Semaglutide CASEIDs
merged_all_q3_ids <- unique(merged_all_q3_clean$caseid)
```

```{r}
# Define severity ranking (custom order)
severity_order <- c("DE", "LT", "HO", "DS", "CA", "RI", "OT")

```

```{r}
# Clean dataset so that each CASEID has only the most serious outcome
merged_all_q3_serious <- merged_all_q3_clean %>%
  mutate(
    severity_rank = match(outc_cod, severity_order)  # assign rank
  ) %>%
  group_by(caseid) %>%
  slice_min(order_by = severity_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-severity_rank)  # drop helper column if not needed
```

```{r}
# Check results, compared raw merged to serious merged, then serious merged to reference unique case id
nrow(merged_all_q3)                     # original rows (with duplicates)
length(unique(merged_all_q3$caseid))    # number of unique CASEIDs
nrow(merged_all_q3_serious)             # should equal number of unique CASEIDs
```

```{r}
# Write the 2024 Q3 Full csv file
fwrite(merged_all_q3_serious, "semaglutide_2024Q3_full.csv")
```

# Q4 2024 DATA MANAGEMENT ONLY

```{r}
# Paths
faers_path4 <- "/Users/mknyarko/Documents/COURSES BROWN SCHOOL/CAPSTONE/CAPSTONE PROJECT/ANALYSIS FILES/FAERS DATA 2024/faers_ascii_2024q4/ASCII"
```

```{r}
# Load DRUG table
drug_q4 <- fread(file.path(faers_path4, "DRUG24Q4.txt"), sep = "$")
```

```{r}
# Extract semaglutide-related reports
semaglutide_q4 <- drug_q4[
  grepl("SEMAGLUTIDE|OZEMPIC|WEGOVY|RYBELSUS", drugname, ignore.case = TRUE) |
  grepl("SEMAGLUTIDE", prod_ai, ignore.case = TRUE)
]
```

```{r}
# Vector to keep only Semaglutide CASEIDs
semaglutide_ids_q4 <- unique(semaglutide_q4$caseid)
```

```{r}
# Load other Q4 tables
demo_q4 <- fread(file.path(faers_path4, "DEMO24Q4.txt"), sep = "$")
reac_q4 <- fread(file.path(faers_path4, "REAC24Q4.txt"), sep = "$")
outc_q4 <- fread(file.path(faers_path4, "OUTC24Q4.txt"), sep = "$")
```

```{r}
# Restrict to semaglutide CASEIDs
  demo_q4_ids <- demo_q4[caseid %in% semaglutide_ids_q4]
  reac_q4_ids <- reac_q4[caseid %in% semaglutide_ids_q4]
  outc_q4_ids <- outc_q4[caseid %in% semaglutide_ids_q4]
```

```{r}
# Limit to U.S. cases only
  demo_q4_ids_US <- demo_q4_ids[reporter_country == "US"]
```

```{r}
# Merge: demographics + outcomes
  merged_demo_outc_q4 <- merge(demo_q4_ids_US, outc_q4_ids, by = "caseid", all.x = TRUE)
```

```{r}
# Merge in reactions
  merged_d_o_reac_q4 <- merge(merged_demo_outc_q4, reac_q4_ids, by = "caseid", all.x = TRUE)
```

# Merge in semaglutide drug info
# merged_all_q4 <- merge(merged_d_o_reac_q4, semaglutide_q4, by = "caseid", all.x = TRUE) 

#First merge attempt with semaglutide_q4 dataset produced an error as there may be duplicate case id's in there
# Semaglutide_q4 dataset was de-duplicated (by selecting first line case id only for every duplicated set)
# Merger was re-attempted

```{r}
# De-duplicating the semaglutide data
semaglutide_q4_unique <- semaglutide_q4[
  !duplicated(caseid), 
]
```

```{r}
# Re-attempt merge in semaglutide drug info
  merged_all_q4 <- merge(merged_d_o_reac_q4, semaglutide_q4_unique, by = "caseid", all.x = TRUE)
```


```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q4)
```


```{r}
# Make duplicate names unique
names(merged_all_q4) <- make.unique(names(merged_all_q4))

```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q4)
```


```{r}
# Remove duplicate columns
merged_all_q4_clean <- merged_all_q4 %>%
  select(-primaryid.x.1, -primaryid.y.1)
```

```{r}
# Checked the actual names in the dataframe, not the names presented in data view
names(merged_all_q4_clean)
```


```{r}
# Vector to keep only Semaglutide CASEIDs
merged_all_q4_ids <- unique(merged_all_q4_clean$caseid)
```

```{r}
# Define severity ranking (custom order)
severity_order <- c("DE", "LT", "HO", "DS", "CA", "RI", "OT")

```

```{r}
# Clean dataset so that each CASEID has only the most serious outcome
merged_all_q4_serious <- merged_all_q4_clean %>%
  mutate(
    severity_rank = match(outc_cod, severity_order)  # assign rank
  ) %>%
  group_by(caseid) %>%
  slice_min(order_by = severity_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-severity_rank)  # drop helper column if not needed
```

```{r}
# Check results, compared raw merged to serious merged, then serious merged to reference unique case id
nrow(merged_all_q4)                     # original rows (with duplicates)
length(unique(merged_all_q4$caseid))    # number of unique CASEIDs
nrow(merged_all_q4_serious)             # should equal number of unique CASEIDs
```

```{r}
# Write the 2024 Q4 Full csv file
fwrite(merged_all_q4_serious, "semaglutide_2024Q4_full.csv")
```

# COMBINING ACROSS ALL 4 QUARTERS

```{r}
# Load all four cleaned quarterly datasets
q1 <- fread("semaglutide_2024Q1_full.csv")
q2 <- fread("semaglutide_2024Q2_full.csv")
q3 <- fread("semaglutide_2024Q3_full.csv")
q4 <- fread("semaglutide_2024Q4_full.csv")
```

```{r}
# Add a quarter column to each dataset to help identify cases by their quarters in the full file
q1[, quarter := "2024Q1"]
q2[, quarter := "2024Q2"]
q3[, quarter := "2024Q3"]
q4[, quarter := "2024Q4"]

```

```{r}
# Now bind them together
semaglutide_2024_all <- rbindlist(
  list(q1, q2, q3, q4), 
  use.names = TRUE, 
  fill = TRUE
)
```

```{r}
# Checked for duplicate columns (variables)
names(semaglutide_2024_all)
```

```{r}
# Checked for duplicate rows
length(unique(semaglutide_2024_all$caseid))    # number of unique CASEIDs
nrow(semaglutide_2024_all)             # should equal number of unique CASEIDs
# If there are duplicate rows, line 1 should yield lower value than line 2
```
# The rows appear duplicated.
# The same caseid can appear in multiple quarters because of case versioning.
# FAERS is cumulative: once a case is reported, it can be updated in later quarters (e.g., new follow-up info, corrected outcome).
# So the same CASEID can appear in Q1 and again in Q3, with updated demographics, drugs, reactions, or outcomes.
# That means this consolidated 2024 dataset may contain duplicates across quarters.

```{r}
# Save consolidated dataset
fwrite(semaglutide_2024_all, "semaglutide_FAERS_2024_full_US.csv")
```

# FURTHER DATA MANAGEMENT TO PREPARE DATA FOR ANALYSIS

```{r}
# Keeping only the most recent version of each case id across quarters
semaglutide_2024_dedup <- semaglutide_2024_all %>%
  group_by(caseid) %>%
  slice_max(order_by = quarter, n = 1, with_ties = FALSE) %>%
  ungroup()
```

```{r}
# Save consolidated deduplicated dataset
fwrite(semaglutide_2024_dedup, "semaglutide_FAERS_2024_full_dedup.csv")
```

```{r}
# Creating subset dataset containing only preferred variables using dplyr
vars <- c("caseid", "caseversion", "age", "age_cod", "sex", 
          "wt", "wt_cod", "reporter_country", "occr_country", "drugname",
           "prod_ai", "dose_amt", "dose_unit", "role_cod", "pt",
           "outc_cod", "event_dt", "rept_dt", "quarter")

semaglutide_2024_subset <- semaglutide_2024_dedup %>%
  select(any_of(vars)) 
```

```{r}
# Creating age group variable inside subset
semaglutide_2024_subset_agegroup <- semaglutide_2024_subset %>%
  mutate(
    age_group = case_when(
      age >= 65 ~ "Older adults (≥65)",
      age >= 18 & age < 65 ~ "Younger adults (18–64)",
      TRUE ~ NA_character_
    )
  ) %>%
  select(
    caseid, caseversion,
    age, age_group, age_cod, sex, wt, wt_cod,
    reporter_country, occr_country,
    drugname, prod_ai, dose_amt, dose_unit, role_cod,
    pt, outc_cod,
    event_dt, rept_dt, quarter
  )
```

```{r}
# Checked for duplicate adverse events
length(unique(semaglutide_2024_subset_agegroup$pt))    # number of unique adverse events
nrow(semaglutide_2024_subset_agegroup)             # should equal number of unique adverse events
# If there are duplicate rows, line 1 should yield lower value than line 2
# Duplicates also signify patterns in ae's
```

# DATA CLEANING FOR NEW SUBSET DATA (WITH AGE GROUPINGS)

```{r}
# Ensuring structure and data types are correct
str(semaglutide_2024_subset_agegroup)
summary(semaglutide_2024_subset_agegroup)
```

```{r}
# Converting variables to factor and numeric as needed
semaglutide_2024_subset_agegroup <- semaglutide_2024_subset_agegroup %>%
  mutate(
    age = as.numeric(age),
    across(c(sex, reporter_country, occr_country, outc_cod, role_cod, quarter),
           ~ as.factor(.))
  )
```

```{r}
summary(semaglutide_2024_subset_agegroup)

fwrite(semaglutide_2024_subset_agegroup, "for_wt_dose_analysis.csv")
```

```{r}
# REMOVING MISSING DATA FROM AGE VARIABLE
semaglutide_2024_age_no_NA <- semaglutide_2024_subset_agegroup %>%
  filter(!is.na(age), age >= 18, age <= 120)

summary(semaglutide_2024_age_no_NA)
```

```{r}
# Replacing all missing spaces with NA in non-numeric (factor) variable columns only

semaglutide_2024_age_no_NA <- semaglutide_2024_age_no_NA %>%
  mutate(
    across(
      where(is.factor),
      ~ {
        # Convert factor → character for cleaning
        x <- as.character(.)
        
        # Trim, collapse extra spaces, and standardize case
        x <- str_squish(str_to_upper(trimws(x)))
        
        # Replace placeholder values with real NA
        x[x %in% c("", "NA", "N/A", "NULL", "NONE", "MISSING", "UNKNOWN", "NOT AVAILABLE")] <- NA
        
        # Convert back to factor
        factor(x)
      }
    )
  )

summary(semaglutide_2024_age_no_NA)
```


```{r}
# Checking for the distinct drugnames
library(dplyr)

semaglutide_2024_age_no_NA %>%
  distinct(drugname)
```

```{r}
# Checking for %proportions of drugnames
semaglutide_2024_age_no_NA %>%
  count(drugname) %>%
  mutate(percentage = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))

```

```{r}
# Standardizing the drugnames
semaglutide_2024_std_drugnames <- semaglutide_2024_age_no_NA %>%
  mutate(
    # Convert everything to uppercase first
    drugname = toupper(drugname),
    
    # Create a clean standardized version
    drug_standardized = case_when(
      #  Combination products FIRST (so they’re not caught by generic SEMAGLUTIDE rule)
      str_detect(drugname, "SEMAGLUTIDE") &
        str_detect(drugname, "CYANOCOBALAMIN|\\+|/|\\\\") ~ "COMBINATION PRODUCT",

      #  Brand-specific products
      str_detect(drugname, "OZEMPIC") ~ "OZEMPIC",
      str_detect(drugname, "WEGOVY") ~ "WEGOVY",
      str_detect(drugname, "RYBELSUS") ~ "RYBELSUS",

      #  Generic semaglutide (unbranded or unknown)
      str_detect(drugname, "SEMAGLUTIDE") &
        !str_detect(drugname, "OZEMPIC|WEGOVY|RYBELSUS") ~ "SEMAGLUTIDE (UNSPECIFIED)",

      #  Everything else
      TRUE ~ "OTHER"
    )
  )%>%                                                   # Rearranging so standardized drug names appear right after drugname variable
  select(
    caseid, caseversion,
    age, age_group, age_cod, sex, wt, wt_cod,
    reporter_country, occr_country,
    drugname, drug_standardized, prod_ai, dose_amt, dose_unit, role_cod,
    pt, outc_cod,
    event_dt, rept_dt, quarter
  )

```

```{r}
# Verifying results
semaglutide_2024_std_drugnames %>%
  count(drug_standardized) %>%
  mutate(percentage = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))

```

```{r}
# Checking for the distinct outcomes
library(dplyr)

semaglutide_2024_std_drugnames %>%
  distinct(outc_cod)

```

```{r}
# Checking for %proportions of outcomes
semaglutide_2024_std_drugnames %>%
  mutate(outc_cod = if_else(is.na(outc_cod), "Missing", outc_cod)) %>%  # Label missing
  count(outc_cod, sort = TRUE) %>%                                     # Count by outcome type
  mutate(
    proportion = round(100 * n / sum(n), 1)                            # Calculate percentage
  )
```

# NA or missiong outcomes make up 50% (>20%) of the data. Cleaning it out will not be recommended for descriptives. Reporting on missing data is a better option. "Not Reported" category to be created. (NB: NR will be cleaned out for regression modelling)
```{r}
# Creating an NR category
semaglutide_2024_outc_NR <- semaglutide_2024_std_drugnames %>%
  mutate(outc_cod = if_else(is.na(outc_cod), "NR", outc_cod))

# All NA's appear as NR (Not reported)

# Verifying results
semaglutide_2024_outc_NR %>%
  count(outc_cod) %>%
  mutate(percentage = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))

```

```{r}
# RE-CONVERTING ALL VARIABLES TO NUMERIC, FACTOR, CHARACTER OR DATE VARIABLES AS NEEDED

semaglutide_2024_outc_NR2 <- semaglutide_2024_outc_NR %>%
  mutate(
    # Identifiers
    caseid = as.character(caseid),               # unique ID, not numeric
    caseversion = as.numeric(caseversion),       # version number for updates
    
    # Demographics
    age = as.numeric(age),                       # continuous variable (years)
    age_group = factor(age_group, 
                       levels = c("YOUNGER ADULTS (18–64)", "OLDER ADULTS (≥65)")),  # categorical
    age_cod = as.character(age_cod),             # usually "YR", "DEC", etc.
    sex = factor(sex, levels = c("M", "F", "UNK")),                    # categorical (male, female, unknown)
    wt = as.numeric(wt),                         # continuous (kg)
    wt_cod = as.character(wt_cod),               # usually "KG" or similar
    
    # Reporter / country
    reporter_country = factor(reporter_country), # categorical (e.g., US)
    occr_country = factor(occr_country),         # location of occurrence
    
    # Drug info
    drugname = factor(drugname),                 # categorical (brand)
    drug_standardized = factor(drug_standardized),      # categorical (brand)
    prod_ai = as.character(prod_ai),             # active ingredient(s)
    dose_amt = as.numeric(dose_amt),             # numeric dose
    dose_unit = as.character(dose_unit),         # dose unit text (mg, ml, etc.)
    role_cod = factor(role_cod, levels = c("PS", "SS", "C", "I")),     # primary/secondary suspect roles
    
    # Event info
    pt = factor(pt),                             # MedDRA-coded preferred term (event)
    outc_cod = factor(outc_cod),                 # outcome codes (DE, HO, OT, etc.)
    event_dt = as.Date(as.character(event_dt), format = "%Y%m%d"),     # date of event
    rept_dt = as.Date(as.character(rept_dt), format = "%Y%m%d"),       # date of report
    quarter = factor(quarter, 
                     levels = c("2024Q1", "2024Q2", "2024Q3", "2024Q4"))  # time period
  )

summary(semaglutide_2024_outc_NR2)
```

```{r}
# Checking proportion of preferred terms (adverse events)

semaglutide_2024_outc_NR2 %>%
  count(pt) %>%
  mutate(percentage = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))

```

```{r}
# Verifying final structure to make sure all age_group and reporter_country categories are valid with no missing data

summary(semaglutide_2024_outc_NR2)
table(semaglutide_2024_outc_NR2$age_group, useNA = "ifany")
table(semaglutide_2024_outc_NR2$reporter_country)
```

```{r}
# Writing the final full dataset

fwrite(semaglutide_2024_outc_NR2, "semaglutide_clean_capstone.csv")
```

```{r}
# Removing non PS role codes so the data retains only Primary Suspect cases
# This will improve causal inference, remove noise and ensure that I am sticking with standard FAERS practice

semaglutide_2024_PS <- semaglutide_2024_outc_NR2 %>%
  filter(toupper(role_cod) == "PS")

summary(semaglutide_2024_PS)
```
# This leaves a total:
# 5302 cases for descriptives and
# 2,604 cases for analysis (Multinomial logistic regression)

```{r}
# Writing the final clean dataset for Capstone

fwrite(semaglutide_2024_PS, "semaglutide_cleanPS_capstone.csv")
```





































































































































































































































